---
title: "Preferences for redistribution in unequal contexts: Changes in Latin America between 2008 and 2018"
subtitle: "Multilevel estimation"
author: "Gonzalo Franetovic - Juan Carlos Castillo"
date: "04-10-2021"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

+ Packages

```{r Packages, message=FALSE, warning=FALSE, include=FALSE}
pacman::p_load(tidyverse, lme4, stats, texreg, car, interplot, sjPlot, 
       data.table, influence.ME, performance, ggeffects, pals, data.table, 
       janitor, xtable, matrixStats, scales, ggthemes)

```

+ Load data 

```{r Database, echo=FALSE, message=FALSE, warning=FALSE}

#Load database
rm(list = ls())
load("Output/LAPOP_Final.RData")

#Rename variables
lapop = lapop %>% 
  mutate(Redistribution = redistribution,
         Income = decile,
         Man = man_f,
         Age = age, 
         Political = ideology_f, 
         Left = left, 
         Married = married_f, 
         SystemConfidence=sysconf, 
         Employment=employment_r, 
         Education=education_r, 
         Urban=zone, 
         Income_Decile=decile_f, 
         GINI=gini,
         GDP=gdp,
         GINI_Mean=gini_mean, 
         GINI_Diff=gini_dif, 
         GDP_Mean=gdp_mean, 
         GDP_Diff=gdp_dif,
         Welfare = welfare,
         Country = countryname,
         CountryWave = country_wave,
         Year = year,
         Decile_d10 = decile_d10,
         Decile_d1 = decile_d1,
         Income2 = Income*Income)

```

# Null model, Cooks Distance and ICC

```{r Null model, echo=FALSE, message=FALSE, warning=FALSE}

#Null model: Country Year (with Venezuela)
model_0_ven = lmer(Redistribution ~ 1 + (1 | CountryWave) + (1 | Country), data=lapop, weights=wt)
screenreg(model_0_ven, stars = c(0.01,0.05,0.1))

#Influence tests
inf.m0 = influence(model_0_ven, "Country")

#Cook D
cooks.distance(inf.m0, parameter=1, sort=TRUE)

jpeg("Output/Graphs/CookD.jpg", width=600, height=400)

plot(inf.m0, which="cook",
cutoff=.17, sort=TRUE,
xlab="Cook´s Distance",
ylab="Country", width=60, height=40)

while (!is.null(dev.list()))  dev.off()

#Exclude Venezuela
lapop = lapop %>% 
  filter(country!="VEN")

country_vars_waves = country_vars_waves %>% 
  filter(country!="VEN")

#Null model: Country Year 
model_0 = lmer(Redistribution ~ 1 + (1 | CountryWave) + (1 |Country), data=lapop, weights=wt)
screenreg(model_0, stars = c(0.01,0.05,0.1))

#Intraclass correlation estimation (ICC)
icc(model_0, by_group = TRUE)

# ICC (Country): 0.015
# ICC (CountryWave): 0.042
# ICC (individual): 0.943

```

# Multilevel estimation

```{r Multilevel estimation, echo=FALSE, message=FALSE, warning=FALSE}

# Model 1. Only income
model_1 = lmer(Redistribution ~ 1 + Income
                + Year + (1 | CountryWave) 
                + (1 | Country), data=lapop, weights=wt)

# Model 2. Individual predictors 
model_2 = lmer(Redistribution ~ 1 + Income + Man + Age + Married 
                + Political + SystemConfidence + Employment + Education 
                + Urban + Year + (1 | CountryWave) + (1 | Country), data=lapop, weights=wt)

# Model 3. Individual predictors + GINI  + GDP
model_3 = lmer(Redistribution ~ 1 + Income + Man + Age + Married 
                + Political + SystemConfidence + Employment + Education 
                + Urban + GINI_Mean + GINI_Diff
                + GDP_Mean + GDP_Diff + Year
                + (1 | CountryWave) + (1 | Country), data=lapop, weights=wt)

# Model 4. Individual predictors + GINI + GDP (Income)
model_4 = lmer(Redistribution ~ 1 + Income + Man + Age + Married 
                + Political + SystemConfidence + Employment + Education 
                + Urban + GINI_Mean + GINI_Diff
                + GDP_Mean + GDP_Diff + Year
                + (1 + Income| CountryWave) + (1 + Income | Country), data=lapop, weights=wt)

#Table
screenreg(list(model_1, model_2, model_3, model_4),  stars = c(0.01,0.05,0.1))

```

# Cross-level interactions

```{r Cross-level interactions, echo=FALSE, message=FALSE, warning=FALSE}

# Model 5: Individual predictors + GINI + GDP (Random Income) + Income*GINI
model_5 = lmer(Redistribution ~ 1 + Income + Man + Age + Married 
                + Political + SystemConfidence + Employment + Education 
                + Urban + GINI_Mean + GINI_Diff
                + GDP_Mean + GDP_Diff
                + Income*GINI_Mean + Income*GINI_Diff + Year
                + (1 + Income | CountryWave) + (1 + Income | Country), data=lapop, weights=wt)

# Model 6: Individual predictors + GINI + GDP (Random Income) + Income*GDP
model_6 = lmer(Redistribution ~ 1 + Income + Man + Age + Married 
                + Political + SystemConfidence + Employment + Education 
                + Urban + GINI_Mean + GINI_Diff
                + GDP_Mean + GDP_Diff
                + Income*GDP_Mean + Income*GDP_Diff + Year
                + (1 + Income | CountryWave) + (1 + Income | Country), data=lapop, weights=wt)

#Table
screenreg(list(model_5, model_6),  stars = c(0.01,0.05,0.1))

```

# Random effect of income

```{r Random effect of income, echo=FALSE, message=FALSE, warning=FALSE}

#Figure 6: Income random effect on preference for redistribution by country: intercept and slope
figure_6 = plot_model(model_4, type = "re", 
                   show.legend = FALSE,
                   show.values = TRUE,
                   facet.grid = FALSE,
                   y.offset = .4,
                   value.offset = .4,
                   value.size = 3.5,
                   sort.est= "Income",
                   title = " ") 

figure_6a = figure_6[[2]] + 
  #scale_y_continuous(limits = c(-0.4, 0.4)) + 
  scale_color_grey(start=0.6, end=0.2) + 
  theme(panel.grid = element_line(colour = "grey"), 
        panel.grid.major.y = element_line(colour = "grey"),panel.background = element_rect(fill = "white")) + 
  geom_hline(yintercept = 0, linetype = "dashed")

figure_6a
ggsave("Output/Graphs/Figure6.png", plot=figure_6a, height = 5, units="in")

ranef(model_6)


#Alternative figure
me = ggpredict(model_4, terms = c("Income", "Country"), type = "re")

figure_6b = plot(me, 
                 show.title = FALSE) +
  labs(
    x = "Income (deciles)", 
    y = "Preference for redistribution", 
    title = " "
  )  +
  scale_x_continuous("Income (deciles)", limits = c(1,10), breaks = c(1,2,3,4,5,6,7,8,9,10)) +
  theme(panel.grid.minor.x = element_blank()) + 
  scale_fill_manual(values=as.vector(polychrome(17)))

figure_6b

# ggsave("../Output/Graphs/Figure_6b.png", plot=figure_6b)

```

# Cross-level interactions: Predicted values

```{r Predicted values, echo=FALSE, message=FALSE, warning=FALSE}

#GINI [WE]
int_inc_ine = ggpredict(model_5, terms = c("Income", "GINI_Diff [-3.97, 4.93]"))

figure_7a = plot(int_inc_ine, 
                 show.title = FALSE, ci.style = "errorbar", dot.size = 2.5) +
  labs(
    x = "Income (deciles)", 
    y = "Preference for redistribution", 
    title = " ",
    colour ="GINI [WE]"
  )  +
  scale_x_continuous("Income (deciles)", limits = c(0.9,10.1), breaks = c(1,2,3,4,5,6,7,8,9,10)) +
  theme(panel.grid.minor.x = element_blank()) +
  scale_y_continuous("Preference for redistribution", limits = c(5, 6.5), breaks = c(5, 5.5, 6, 6.5)) +
  scale_colour_grey(start=0.6, end=0.1, name="GINI [WE]", labels = c("Max decrease", "Max increase"))

figure_7a
ggsave("Output/Graphs/Figure7a.png", plot=figure_7a, height = 5, units="in")

#GDP [BE]
int_inc_dev = ggpredict(model_6, terms = c("Income", "GDP_Mean [1.65, 13.98]"))

figure_7b = plot(int_inc_dev, 
                 show.title = FALSE, ci.style = "errorbar", dot.size = 1.5) +
  labs(
    x = "Income (deciles)", 
    y = "Preference for redistribution", 
    title = " ",
    colour ="GDP [BE]"
  )  +
  scale_x_continuous("Income (deciles)", limits = c(0.9,10.1), breaks = c(1,2,3,4,5,6,7,8,9,10)) +
  theme(panel.grid.minor.x = element_blank())  +
  scale_y_continuous("Preference for redistribution", limits = c(5, 6.5), breaks = c(5, 5.5, 6, 6.5)) +
  scale_colour_grey(start=0.6, end=0.1, name="GDP [BE]", labels = c("Low", "High"))

figure_7b
ggsave("Output/Graphs/Figure7b.png", plot=figure_7b, height = 5, units="in")

```

# Appendix: Tables and complementary figures

+ Appendix 1 - Sample: Observations by country and year

```{r Appendix 1, echo=FALSE, message=FALSE, warning=FALSE}

#Table_ Sample: Observations by country and Year
lapop$cases=1
sample = with(lapop, tapply(cases, list(Country, Year), sum))
sample = as.data.frame(sample)
sample$Total = rowSums(sample[c("2008","2010","2012", "2014", "2016", "2018")])

sample_country = as.data.frame(with(lapop, tapply(cases, list(Year), sum)))
colnames(sample_country) = "Total"
sample_country$Total = as.numeric(sample_country$Total)
setDT(sample_country, keep.rownames = TRUE)[]
sample_country2 = data.frame(t(sample_country)) %>% 
  row_to_names(row_number = 1)
sample_country2[] = lapply(sample_country2, function(x) as.numeric(as.character(x)))
sample_country2$Total = rowSums(sample_country2)
sample1 = rbind(sample, sample_country2)

xtable(sample1, digits = 0)

```

+ Appendix 2 - Within and between-country distribution of dependent and independent variables.

```{r Appendix 2, echo=FALSE, message=FALSE, warning=FALSE}

#Table: Within- and between-country distribution of dependent and independent variables

##Preference for redistribution

#Country names
lapop$countryname1 = as.factor(lapop$countryname)
table(lapop$countryname1)

#Mean by country
redis_country_mean = with(lapop, tapply(redistribution, list(countryname1), mean))
redis_country_mean = as.data.frame(redis_country_mean)
redis_country_mean0 = as.data.frame(redis_country_mean[ order(row.names(redis_country_mean)), ])

#Sd by country
redis_country_sd = with(lapop, tapply(redistribution, list(countryname1), sd))
redis_country_sd = as.data.frame(redis_country_sd)

#Sd by country-year
redis_country_year = with(lapop, tapply(redistribution, list(countryname1, year), mean))
redis_country_year = as.data.frame(redis_country_year)

redis_country_year$total = rowMeans(redis_country_year[c("2008", "2010", "2012", "2014", "2016", "2018")], na.rm=TRUE)

redis_country_year1 = redis_country_year
setDT(redis_country_year1, keep.rownames = TRUE)[]
redis_country_year1 = redis_country_year1 %>% 
  arrange(rn)

redis_country_sdyear = as.matrix(redis_country_year)
redis_country_sdyear = rowSds(redis_country_sdyear, na.rm=TRUE)


##Country level variables

lapop_names = lapop %>% 
  distinct(country, countryname1)

country_vars_waves = country_vars_waves %>% 
  mutate(year = as.factor(year))

country_vars_waves_name = left_join(lapop_names, country_vars_waves)

##GINI

#Mean by country
gini_country_mean = with(country_vars_waves_name, tapply(gini, list(countryname1), mean))
gini_country_mean = as.data.frame(gini_country_mean)

#Sd by country-year
gini_country_year = with(country_vars_waves_name, tapply(gini, list(countryname1, year), mean))
gini_country_year = as.data.frame(gini_country_year)

gini_country_year$total = rowMeans(gini_country_year[c("2008", "2010", "2012", "2014", "2016", "2018")], na.rm=TRUE)

gini_country_sdyear = as.matrix(gini_country_year)

gini_country_sdyear= rowSds(gini_country_sdyear, na.rm=TRUE)

##GDP

#Mean by country
gdp_country_mean = with(country_vars_waves_name, tapply(gdp, list(countryname1), mean))
gdp_country_mean = as.data.frame(gdp_country_mean)

#Sd by country-year
gdp_country_year = with(country_vars_waves_name, tapply(gdp, list(countryname1, year), mean))
gdp_country_year = as.data.frame(gdp_country_year)

gdp_country_year$total = rowMeans(gdp_country_year[c("2008", "2010", "2012", "2014", "2016", "2018")], na.rm=TRUE)

gdp_country_sdyear= as.matrix(gdp_country_year)

gdp_country_sdyear = rowSds(gdp_country_sdyear, na.rm=TRUE)


##Total
redis_mean = round(mean(lapop$redistribution), 3)
redis_sd = round(sd(lapop$redistribution), 3)

gini_mean = round(mean(country_vars_waves$gini), 3)
gini_sd = round(sd(country_vars_waves$gini), 3)

gdp_mean = round(mean(country_vars_waves$gdp), 3)
gdp_sd = round(sd(country_vars_waves$gdp), 3)

annex2_total = c("Total", redis_mean, redis_sd, "", gini_mean, gini_sd, gdp_mean, gdp_sd)


##Table
annex2a = as.data.frame(cbind(redis_country_mean, redis_country_sd, redis_country_sdyear,
                             gini_country_mean, gini_country_sdyear,
                             gdp_country_mean, gdp_country_sdyear))

setDT(annex2a, keep.rownames = TRUE)[]

annex2a = annex2a %>% 
  rename(countryname = rn) %>% 
  arrange(countryname)

annex2_totals = c("Total", redis_mean, redis_sd, NA, gini_mean, gini_sd, gdp_mean, gdp_sd)
annex2_totals = as.data.frame(t(annex2_totals))

#Add totals
annex2 = rbind(annex2a, annex2_totals, use.names=FALSE)

#Round numbers
indexes = annex2 %>% 
  dplyr::select(-countryname)

countries = annex2 %>% 
  dplyr::select(countryname)

indexes[] = lapply(indexes, function(x) as.numeric(as.character(x)))
indexes[] = lapply(indexes, function(x) if(is.numeric(x)) round(x, 3) else x)

annex2c = cbind(countries, indexes)

#Final table
annex2 = xtable(annex2c, digits = 3)
annex2

names(annex2) = c("Country", "Redist Mean", "Redist SD total", "Redist SD years", "Gini Mean", "Gini SD years", "GDP Mean", "GDP SD years")
print(xtable(annex2), include.rownames=FALSE, display = c("s","fg","d","d","d","d","d","d"))

```

+ Appendix 3 - Average preference for redistribution, by country and year

```{r Appendix 3, echo=FALSE, message=FALSE, warning=FALSE}

#Preference for redistribution by country and year

redis_country_year = with(lapop, tapply(redistribution, list(countryname, year), mean))
redis_country_year = as.data.frame(redis_country_year)
redis_country_year = as.data.frame(setDT(redis_country_year, keep.rownames = TRUE)[])
redis_country_year = melt(redis_country_year, id=c("rn"))

appendix3 = ggplot(redis_country_year, aes(variable, value))+
  geom_bar(stat = "identity", fill = "#333333") +
    labs(x = "Year", y =  "Preference for redistribution") + 
  facet_wrap(~ rn, ncol = 3)  +
  scale_y_continuous(limits=c(1,7), breaks = c(1,2,3,4,5,6,7), oob = rescale_none) + 
  theme(
      axis.text=element_text(size=13),
          strip.text=element_text(size=15),
          legend.text=element_text(size=15),
          legend.title=element_text(size=15),
          legend.key.size=unit(0.5,"cm"),
          legend.position = "bottom",
          panel.background = element_rect(fill = "white")) +
    theme_hc()  

appendix3

ggsave("Output/Graphs/Appendix3.png", plot=appendix3, height = 10, width = 9, units="in")
```

+ Appendix 4 - Hybrid multilevel regression models of individual preference for redistribution. Continuous,categorical and quadratic income measures.

```{r Appendix 4, echo=FALSE, message=FALSE, warning=FALSE}

#Lineal
model_2 = lme4::lmer(Redistribution ~ 1 + Income + Man + Age + Married 
                + Political + SystemConfidence + Employment + Education 
                + Urban + Year + (1 | CountryWave) + (1 | Country), data=lapop, weights=wt)


#Factor
model_2_f = lme4::lmer(Redistribution ~ 1 + Income_Decile + Man + Age + Married 
                + Political + SystemConfidence + Employment + Education 
                + Urban + Year + (1 | CountryWave) + (1 | Country), data=lapop, weights=wt)


#Quadratic
model_2_q = lme4::lmer(Redistribution ~ 1 + Income + I(Income^2) + Man + Age + Married 
                + Political + SystemConfidence + Employment + Education 
                + Urban + Year + (1 | CountryWave) + (1 | Country), data=lapop, weights=wt)

#Table
screenreg(list(model_2, model_2_f, model_2_q),  stars = c(0.01,0.05,0.1))

```


# Extra

+ Model 3 with welfare state typology

```{r Extra 1, echo=FALSE, message=FALSE, warning=FALSE}

#Model 7: Individual predictors + GINI  + GDP + Welfare State
model_7 = lmer(Redistribution ~ 1 + Income + Man + Age + Married 
                + Political + SystemConfidence + Employment + Education 
                + Urban + GINI_Mean + GINI_Diff
                + GDP_Mean + GDP_Diff
                + Welfare + Year
                + (1 | CountryWave) + (1 | Country), data=lapop, weights=wt)

screenreg(list(model_3, model_7),  stars = c(0.01,0.05,0.1))

```

+ Relation between and within countries

```{r Extra 2, echo=FALSE, message=FALSE, warning=FALSE}

between = lapop %>% 
  mutate(wave = as.numeric(paste0(wave))) %>% 
  group_by(country) %>% 
  summarise_all(mean, na.rm=TRUE)

within = lapop %>% 
  mutate(wave = as.numeric(paste0(wave))) %>% 
  group_by(countryname1, Year) %>% 
  summarise_all(mean, na.rm=TRUE)

extra1= ggplot(between, 
                 aes(x= GINI_Mean, y= Redistribution, label=country)) +
  geom_point() + 
  geom_smooth(method = "lm", colour="black") +
  geom_text(aes(label=country),hjust=-0.1, vjust=-0.5, size=3.5) +
  labs(x = "Inequality", y = "Redistributive Preference", fill="") +
  scale_y_continuous("Redistributive Preference", limits = c(4.8,6.2), 
                     breaks = c(4.8,5.2,5.6,6)) +
  scale_x_continuous("Inequality", limits = c(40,60),
                     breaks = c(40,50,60)) +    
  theme(panel.grid.major = element_line(colour = "grey"),
        legend.position="none",
        axis.text=element_text(size=10),
        strip.text=element_text(size=12),
        legend.text=element_text(size=12),
        legend.title=element_text(size=12),
        legend.key.size=unit(1,"cm"),
        panel.background = element_rect(fill = "white"))
extra1
ggsave("Output/Graphs/Extra1.png", plot=extra1)

#Inequality: Relation "within" countries
extra2= ggplot(within, 
                 aes(x= GINI, y= Redistribution, colour=as.factor(wave))) +
  geom_point(size = 2, alpha = .8) + 
  stat_smooth(size = 1, method = "lm", se = FALSE, colour="black") +
  facet_wrap(~countryname1, ncol=3) +
  labs(x = "Inequality", y =  "Redistributive Preference") + 
  scale_color_grey(start=0.8, end=0.2, name="Wave") +  
  scale_y_continuous( "Redistributive Preference", limits = c(4.2,6.7), 
                      breaks = c(4.5,5,5.5,6,6.5)) +
  scale_x_continuous("Inequality",
                     breaks = c(35,40,45,50,55,60)) +  
  theme(panel.grid.major = element_line(colour = "grey"),
        legend.position="bottom",
        axis.text=element_text(size=10),
        strip.text=element_text(size=12),
        legend.text=element_text(size=12),
        legend.title=element_text(size=12),
        legend.key.size=unit(1,"cm"),
        panel.background = element_rect(fill = "white"))
extra2
ggsave("Output/Graphs/Extra2.png", plot=extra2, height = 10, width = 10, units="in")

#Economic Development: Relation "between" countries
extra3 = ggplot(between, 
                  aes(x= GDP_Mean, y= Redistribution, label=country)) +
  geom_point() + 
  geom_smooth(method = "lm", colour="black") +
  geom_text(aes(label=country),hjust=-0.1, vjust=-0.5, size=3.5) +
  labs(x = "Economic Development", y = "Redistributive Preference", fill="") +
  scale_y_continuous("Redistributive Preference", limits = c(4.8,6.2), 
                     breaks = c(4.8,5.2,5.6,6)) +
  scale_x_continuous("Economic Development", limits = c(1,14),
                     breaks = c(5,10,15)) +  
  theme(panel.grid.major = element_line(colour = "grey"),
        legend.position="none",
        axis.text=element_text(size=10),
        strip.text=element_text(size=12),
        legend.text=element_text(size=12),
        legend.title=element_text(size=12),
        legend.key.size=unit(1,"cm"),
        panel.background = element_rect(fill = "white"))
extra3
ggsave("Output/Graphs/Extra3.png", plot=extra3)

#Economic Development: Relation "within" countries
extra4= ggplot(within, 
                 aes(x= GDP, y= Redistribution, colour=as.factor(wave))) +
  geom_point(size = 2, alpha = .8) + 
  stat_smooth(size = 1, method = "lm", se = FALSE, colour="black") +
  facet_wrap(~countryname1, ncol=3) +
  labs(x = "Economic Development", y =  "Redistributive Preference") + 
  scale_color_grey(start=0.8, end=0.2, name="Wave") +  
  scale_y_continuous( "Redistributive Preference", limits = c(4.2,6.7), 
                      breaks = c(4.5,5,5.5,6,6.5)) +
  scale_x_continuous("Economic Development",
                     breaks = c(0,5,10,15,20)) +  
  theme(panel.grid.major = element_line(colour = "grey"),
        legend.position="bottom",
        axis.text=element_text(size=10),
        strip.text=element_text(size=12),
        legend.text=element_text(size=12),
        legend.title=element_text(size=12),
        legend.key.size=unit(1,"cm"),
        panel.background = element_rect(fill = "white"))
extra4
ggsave("Output/Graphs/Extra4.png", plot=extra4, height = 10, width = 10, units="in")
```


