---
title: "Preferences for redistribution in unequal contexts: Changes in Latin America between 2008 and 2018"
subtitle: "Descriptive statistics"
author: "Gonzalo Franetovic - Juan Carlos Castillo"
date: "04-10-2021"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

+ Packages

```{r Packages, message=FALSE, warning=FALSE, include=FALSE}

#Install and load packages
pacman::p_load(tidyverse, countrycode, car, zoo, wbstats, Hmisc, stargazer, stats,
       ggthemes, data.table, reshape2, scales, xtable, ggrepel)

```

+ Load data 

```{r Database, echo=FALSE, message=FALSE, warning=FALSE}
#Load database
rm(list = ls())
load("Input/proc/LAPOP_Final.RData")

#Exclude Venezuela
lapop_ven = lapop
country_vars_waves_ven = country_vars_waves

lapop = lapop %>% 
  filter(country!="VEN")

country_vars_waves = country_vars_waves %>% 
  filter(country!="VEN")

```

# Descriptive statistics 

```{r Descriptive statistics, echo=FALSE, message=FALSE, warning=FALSE}
#New 
lapop = lapop %>% 
  select(country, countryname, country_wave, country_wave_rec, 
         year, wave, weight1500, wt, redistribution, 
         man_f, age, married_f, ideology_f, employment_r, education_r, 
         decile, decile_d10, decile_f, zone, sysconf, 
         gdp, gdp_mean, gdp_dif, gini, gini_mean, gini_dif, welfare)

#Level 1
lapop_a = lapop %>% 
  select(redistribution, decile, man_f, age, 
         married_f, ideology_f, sysconf, employment_r, 
         education_r, zone, decile_d10, decile_f) 

lapop_a = lapop_a %>%
  rename(income = decile) %>% 
  mutate(man = as.numeric(man_f) - 1,
         married = as.numeric(married_f) - 1,
         ideology = as.numeric(ideology_f) - 1,
         urban = as.numeric(zone) - 1,
         employment = as.numeric(employment_r) - 1,
         education = as.numeric(education_r) - 1)


stargazer(lapop_a,title="Descriptive Statistics", omit.summary.stat = c("p25", "p75"))

prop.table(table(lapop_a$man_f))
prop.table(table(lapop_a$married_f))
prop.table(table(lapop_a$employment_r))
prop.table(table(lapop_a$ideology_f))
prop.table(table(lapop_a$education_r))
prop.table(table(lapop_a$zone))

#Level 2
country_vars_waves_a= country_vars_waves %>% 
  select(-c(year, welfare))

stargazer(country_vars_waves_a, title="Descriptive Statistics", omit.summary.stat = c("p25", "p75"))

#Level 3
country_vars = country_vars_waves %>% 
  distinct(country, .keep_all=TRUE) %>% 
  select(country, welfare)

prop.table(table(country_vars$welfare))

```

# Preference for redistribution by countries. Percentage by category.

```{r Figure 2, echo=FALSE, message=FALSE, warning=FALSE}

#Figure 2: Preference for redistribution by countries. Percentage by category.
redist = xtabs(~countryname+redistribution, data = lapop)
redist = prop.table(redist, 1)*100
redist = as.data.frame(redist) %>% 
  arrange(desc(redistribution, Freq)) %>% 
  mutate(countryname1 = case_when(countryname=="Bolivia" ~ 1,
                                  countryname=="Peru" ~ 2,
                                  countryname=="Guatemala" ~ 3,
                                  countryname=="Ecuador" ~ 4,
                                  countryname=="Honduras" ~ 5,
                                  countryname=="El Salvador" ~ 6,
                                  countryname=="Mexico" ~ 7,
                                  countryname=="Colombia" ~ 8,
                                  countryname=="Panama" ~  9,
                                  countryname=="Brazil" ~ 10,
                                  countryname=="Paraguay" ~ 11,
                                  countryname=="Uruguay" ~ 12,
                                  countryname=="Chile" ~ 13,
                                  countryname=="Argentina" ~ 14,
                                  countryname=="Costa Rica" ~ 15,
                                  countryname=="Dominican Republic" ~ 16,
                                  countryname=="Nicaragua" ~ 17)) %>% 
  mutate(countryname1 = factor(countryname1,
         levels = c(1:17),
         labels = c("Bolivia", "Peru", "Guatemala", "Ecuador", "Honduras", "El Salvador", "Mexico",
                    "Colombia", "Panama", "Brazil", "Paraguay", "Uruguay", "Chile", "Argentina", 
                    "Costa Rica", "Dominican Republic", "Nicaragua")),
         order = factor(countryname,
         levels = c(1:17),
         labels = c("Bolivia", "Peru", "Guatemala", "Ecuador", "Honduras", "El Salvador", "Mexico",
                    "Colombia", "Panama", "Brazil", "Paraguay", "Uruguay", "Chile", "Argentina", 
                    "Costa Rica", "Dominican Republic", "Nicaragua")))

  figure2 = ggplot(data = redist, 
                   aes(x=reorder(countryname1, order), y = Freq, 
                       fill = factor(redistribution))) + 
    geom_bar(stat = "identity", position = position_stack(reverse = TRUE)) + coord_flip() + labs(x = "", y = "", fill="") +
    guides(fill = guide_legend(reverse = TRUE)) + 
    scale_y_continuous(labels = function(Freq){paste0(Freq, "%")}) +
    theme(panel.background = element_rect(fill = "white"), 
          axis.text=element_text(size=10),
          strip.text=element_text(size=12))  + 
    scale_fill_grey(start=0.8, end=0.2, name = "", 
                  labels = c("1 (Disagree)", "2", "3", "4", "5", "6", "7 (Agree)")) 

figure2
ggsave("Output/Graphs/Figure2.png", plot=figure2, height = 4, units="in")

```
# Preference for redistribution by country and year. Percentage by category.

```{r Figure 3, echo=FALSE, message=FALSE, warning=FALSE}

#Figure 2: Preference for redistribution by country and year
figure3 = ggplot(lapop, aes(x= redistribution,  colour= year)) + 
  geom_line(aes(y = ..prop.., fill = factor(..x..)), stat="count", binwidth = 1, size = 1) +
  xlim(1, 7) +
  facet_wrap(~countryname, ncol=3) +
  labs(x = "Preference for redistribution", y = "", fill="") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_x_continuous(breaks = c(1:7)) +
  theme(
    axis.text=element_text(size=15),
        strip.text=element_text(size=15),
        legend.text=element_text(size=15),
        legend.title=element_text(size=15),
        legend.key.size=unit(0.5,"cm"),
        legend.position = "bottom",
        panel.background = element_rect(fill = "white")) +
  theme_hc() + scale_color_grey(start=0.9, end=0.1, name="Wave") 

figure3
ggsave("Output/Graphs/Figure3.png", plot=figure3, height = 10, width = 9, units="in")

```

+ Average preference for redistribution of income deciles, by country.

```{r Country and income, echo=FALSE, message=FALSE, warning=FALSE}

#Table 3: Average redistribution preference of income deciles, by country
redis_decile = with(lapop, tapply(redistribution, list(countryname, decile_f), mean))
redis_decile = as.data.frame(redis_decile)
redis_decile = as.data.frame(setDT(redis_decile, keep.rownames = TRUE)[])

redis_decile_country = with(lapop, tapply(redistribution, list(countryname), mean))
redis_decile_country = as.data.frame(redis_decile_country)
redis_decile_country = as.data.frame(setDT(redis_decile_country, keep.rownames = TRUE)[])
colnames(redis_decile_country) = c("rn", "Total")

redis_decile = left_join(redis_decile, redis_decile_country) %>% 
  rename(country = rn)

redis_decile

print(xtable(redis_decile), include.rownames=FALSE)

```

# Average preference for redistribution and GINI, by country and year.

```{r Figure 4, echo=FALSE, message=FALSE, warning=FALSE}

#Within countries
within = lapop %>% 
  mutate(wave = as.numeric(paste0(wave))) %>% 
  group_by(country, wave) %>% 
  summarise_all(mean, na.rm=TRUE)

#Figure 4: Average redistribution preference and inequality, by country and year
figure4= ggplot(within, 
            aes(x= gini, y= redistribution, label=country)) +
  geom_point(size = 1.5, alpha = 0.8) + 
  stat_smooth(size = 1, method = "lm", se = FALSE, fullrange = TRUE, colour = "black") +   
  geom_text_repel(aes(label=country), size=2.5, show.legend = FALSE) +
  labs(x = "Inequality", y =  "Preference for redistribution") + 
  scale_color_grey(start=0.8, end=0.2, name="Wave") +  
  scale_y_continuous( "Average preference for redistribution", limits = c(4.5,6.6), 
                     breaks = c(4.5,5, 5.5, 6, 6.5)) +
  scale_x_continuous("GINI", limits = c(39,60), 
                     breaks = c(40,45,50,55,60)) +
  theme(panel.grid.major = element_line(colour = "grey"),
        legend.position="bottom",
        axis.text=element_text(size=10),
        strip.text=element_text(size=10),
        legend.text=element_text(size=10),
        legend.title=element_text(size=10),
        legend.key.size=unit(1,"cm"),
        panel.background = element_rect(fill = "white")) +  
  facet_wrap(~ wave) 

figure4
ggsave("Output/Graphs/Figure4.png", plot=figure4, height = 5, units="in")

```

# Average preference for redistribution and GDP per capita, by country and year.

```{r Figure 5, echo=FALSE, message=FALSE, warning=FALSE}

#Figure 5: Average redistribution preference and economic development, by country and year (facet)
figure5= ggplot(within, 
            aes(x= gdp, y= redistribution, label=country)) +
  geom_point(size = 1.5, alpha = 0.8) + 
  stat_smooth(size = 1, method = "lm", se = FALSE, fullrange = TRUE, colour = "black") +   
  geom_text_repel(aes(label=country), size=2.5, show.legend = FALSE) +
  labs(x = "Inequality", y =  "Preference for redistribution") + 
  scale_color_grey(start=0.8, end=0.2, name="Wave") +  
  scale_y_continuous( "Average preference for redistribution", limits = c(4.5,6.6), 
                     breaks = c(4.5,5, 5.5, 6, 6.5)) +
  scale_x_continuous("GDP per capita", limits = c(0,16), 
                     breaks = c(0,4,8,12,16)) +
  theme(panel.grid.major = element_line(colour = "grey"),
        legend.position="bottom",
        axis.text=element_text(size=10),
        strip.text=element_text(size=10),
        legend.text=element_text(size=10),
        legend.title=element_text(size=10),
        legend.key.size=unit(1,"cm"),
        panel.background = element_rect(fill = "white")) +  
  facet_wrap(~ wave) 

figure5
ggsave("Output/Graphs/Figure5.png", plot=figure5, height = 5, units="in")

```


