---
title: "Preferences for income redistribution in unequal contexts: Changes in Latin America between 2008 and 2018"
subtitle: "Database processing"
author: "Gonzalo Franetovic - Juan Carlos Castillo"
date: "04-10-2021"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

The research code document is detailed below. The adjustment of the databases, cleaning and recoding of variables, estimation of models, tables and graphic figures are explained step by step. Other analyzes available in the document "Code annex".

+ Packages

```{r Packages, message=FALSE, warning=FALSE, include=FALSE}

#Install and load packages
#install.packages("pacman")
pacman::p_load(tidyverse, countrycode, car, zoo, wbstats, Hmisc, here, conflicted)

#Packages conflicts
conflict_prefer("select", "dplyr")
conflict_prefer("filter", "dplyr")
conflict_prefer("recode", "car")

```

# Individual level variables

+ Full LAPOP 2008-2018 database paid for by the Institute of Political Science of the Pontificia Universidad Católica de Chile (incorporates information omitted in the free version)

```{r Database Lapop Full, echo=TRUE, message=FALSE, warning=FALSE}
#Load database
rm(list = ls())
load(here("Input", "Original", "2008-2018_LAPOP_AL.RData"))
```

```{r Database Lapop Selection, echo=TRUE, message=FALSE, warning=FALSE}
#lapop = as.data.frame(lapop) #Transformation to Data Frame
names(lapop) = tolower(names(lapop)) #Variable names in lower case

#Name of the variables
lapop = lapop %>%
  select(pais, wave, weight1500, wt, ros4, q1, q2, q11, q11n,
         l1, ocup4a, colocup4a, ed, q10, q10new_12, q10new_14, q10new_16, q10new_18,
         ur, tamano, b12, b18, b31, b21a, b13, b21,
         q12, q12c)

colnames(lapop) = c("country", "year", "weight1500", "wt", "redistribution",
                    "male", "age", "married1", "married2",
                    "ideology", "employment1", "employment2", "education", 
                    "income0810", "income12", "income14", "income16", "income18",
                    "zone", "sizecity",
                    "ffaaconfidence", "policeconfidence", "judicialconfidence",
                    "executiveconfidence", "congressconfidence", "polpartiesconfidence",
                    "children", "housesize")
```

+ Selection and identification of the countries and waves of the sample

```{r Sample of countries and waves, echo=TRUE, message=FALSE, warning=FALSE}
#Remove countries that we will not consider in the sample
lapop = lapop %>% 
  mutate(country = as.numeric(country)) %>% 
  filter(country<=21) %>% 
  mutate(country = ifelse(country==21, 18, country))

#Countries of the sample (Labels)
lapop$country = factor(lapop$country,
  levels = c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18),
  labels = c("MEX", "GTM", "SLV", "HND", "NIC", "CRI", 
             "PAN", "COL", "ECU", "BOL", "PER", "PRY", 
             "CHL", "URY", "BRA", "VEN", "ARG", "DOM"))

lapop$countryname = countrycode(lapop$country,"iso3c","country.name")
lapop$countryname_es = countrycode(lapop$country,"iso3c","un.name.es")

#Remove waves that we will not consider in the sample
lapop = lapop %>% 
  filter(year>=2008)

describe(lapop$country)
describe(lapop$year)
describe(lapop$countryname)
describe(lapop$countryname_es)
```

+ Adjustment of different variables between years

```{r Adjustment of different variables between years, echo=TRUE, message=FALSE, warning=FALSE}
#Adjustment of different variables between years

#Employment
lapop = lapop %>% 
  mutate(employment2 = ifelse(employment2==7, 6, employment2)) %>% 
  mutate(employment2 = ifelse(employment2==8, 7, employment2)) %>%
  mutate(employment = ifelse(country=="COL" & year==2018, employment2, employment1))

lapop = lapop %>% 
  select(-c(employment1, employment2))

#Married
lapop = lapop %>% 
  mutate(married = ifelse(year<=2012, married1, married2))

lapop = lapop %>% 
  select(-c(married1, married2))
```

+ Generation of individual level variables

```{r Generation of individual level variables, echo=TRUE, message=FALSE, warning=FALSE}
#Waves
lapop$wave = lapop$year
lapop$wave = as.factor(lapop$wave)

#Country_year
lapop$country_wave = do.call(paste, c(lapop[c("country", "wave")], sep = "_"))

#Preference for redistribution
#Dummy
lapop$redistribution_r = as.numeric(lapop$redistribution >= 7)
lapop$redistribution_r = as.factor(lapop$redistribution_r)
#Log Natural
lapop$redistribution_ln = log(lapop$redistribution)

#Man
lapop$male = as.numeric(lapop$male <= 1)
lapop$male_f = factor(lapop$male, levels = c(0,1), labels = c("Female", "Male"))

#Age
lapop$age[lapop$age<=17] = NA #Continue Variable
lapop$age2 = (lapop$age)^2 #Quadratic Variable

#Married (or cohabiting)
lapop$married = as.factor(lapop$married)
lapop = lapop %>% 
  mutate(married_f = recode(married,"2=1; 3=1; 7=1; 1=0; 4=0; 5=0; 6=0"))
lapop$married_f = factor(lapop$married_f, levels = c(0,1), labels = c("Single", "Married"))

#Left political ideology
#Continue Variable
lapop$left = 1+10-(lapop$ideology)
#Factor Variable
lapop = lapop %>% 
  mutate(ideology_f = recode(left,"1:4=1; 5:6=2; 7:10=3"))
lapop$ideology_f[is.na(lapop$ideology_f)] = 99
lapop$ideology_f = factor(lapop$ideology_f, levels = c(1,2,3, 99), labels = c("Right", "Center", "Left", "Not declared"))

#Employment
lapop = lapop %>% 
  mutate(employment_r = recode(employment,"4=1; 6=1; 3=2; 5=2; 7=2; 1:2=3"))
lapop$employment_r = factor(lapop$employment_r, levels = c(1,2,3), labels = c("No workforce", "Unemployed", "Employed"))

#Education
lapop = lapop %>% 
  mutate(education_r = recode(education,"0:6=1; 7:12=2; 13:18=3"))
lapop$education_r = factor(lapop$education_r, levels = c(1,2,3), labels = c("Primary", "Secondary", "Tertiary"))

#Urban zone
lapop$zone[lapop$zone==2] = 0
lapop$zone =factor(lapop$zone, levels = c(0, 1), labels = c("Rural", "Urban"))

#City size
lapop$tamano = lapop$sizecity
lapop$sizecity = 1+5-(lapop$tamano)
lapop$sizecity[lapop$sizecity<1] = NA
lapop$sizecity = factor(lapop$sizecity, levels = c(1,2,3,4,5), labels = c("Rural area", "Small city", "Medium city", "Big city", "National capital"))

#Political system confidence
lapop$sysconf = rowMeans(lapop[c("ffaaconfidence", "policeconfidence", "judicialconfidence", "executiveconfidence", "congressconfidence", "polpartiesconfidence")], na.rm=TRUE)

#People in the home: the number of children in the household is taken as a proxy of the variable "people in the home" with a adjust of +1. This is used to estimate per capita income.
lapop$nhome = lapop$children+1

#Filter
lapop = lapop %>% 
  filter(year==2008 | year==2010 | year==2012 | year==2014 | year==2016 | year==2018)
```

```{r Income, echo=TRUE, message=FALSE, warning=FALSE}
#Income
lapop = lapop %>%
  mutate(income_ori = case_when(year<=2010 ~ income0810,
                                year==2012 ~ income12,
                                year==2014 ~ income14,
                                year==2016 ~ income16,
                                year==2018 ~ income18))

#Missings
lapop = lapop %>% 
  mutate(income_ori = ifelse(income_ori==0, NA, income_ori))

#Deciles 2012 - 2018
lapop = lapop %>%
  group_by(country, year) %>%
  mutate(decile=ntile(income_ori,10))

lapop = lapop %>% 
  mutate(income = ifelse(year>=2012, decile, income_ori),
         decile = income)

#Factors

#Decile incorporating missing category
lapop$decile_f = lapop$decile
lapop$decile_f = as.factor(lapop$decile_f)

#Dummies
table(lapop$decile_d10)

#Decile 10 dummy with missing as reference
lapop = lapop %>% 
  mutate(decile_d10 = recode(decile,"1:9=0; 10=1"))
lapop$decile_d10 =factor(lapop$decile_d10, levels = c(0,1), labels = c("Not decile 10", "Decile 10"))
describe(lapop$decile_d10)

#Decile 1 dummy with missing as reference
lapop = lapop %>% 
  mutate(decile_d1 = recode(decile,"2:10=0; 1=1"))
lapop$decile_d1 =factor(lapop$decile_d1, levels = c(0,1), labels = c("Not decile 1", "Decile 1"))
describe(lapop$decile_d1)

#Cleaning database
lapop = lapop %>% 
  select(-c(income0810, income12, income14, income16, income18, income_ori))
```

+ Missing values: Only variables with complete information

```{r Missing values, echo=TRUE, message=FALSE, warning=FALSE}
#Listwise with the variables used to estimate the models
lapop = lapop[complete.cases(lapop$redistribution, lapop$male_f, lapop$age, lapop$married_f, lapop$ideology_f, lapop$employment_r, lapop$education_r, lapop$income, lapop$zone, lapop$sysconf),]
```

+ Time dummy

```{r Time dummy, echo=TRUE, message=FALSE, warning=FALSE}
#2008
lapop$year_2008 = 0
lapop$year_2008[lapop$year==2008] = 1
#2008
lapop$year_2010 = 0
lapop$year_2010[lapop$year==2010] = 1
#2008
lapop$year_2012 = 0
lapop$year_2012[lapop$year==2012] = 1
#2008
lapop$year_2014 = 0
lapop$year_2014[lapop$year==2014] = 1
#2008
lapop$year_2016 = 0
lapop$year_2016[lapop$year==2016] = 1
#2008
lapop$year_2018 = 0
lapop$year_2018[lapop$year==2018] = 1
```

# Country level variables 

```{r Country level variables, echo=TRUE, message=FALSE, warning=FALSE}
#Countries
countries = c("ARG",
              "BOL",
              "BRA",
              "CHL",
              "COL",
              "CRI",
              "DOM",
              "ECU",
              "GTM",
              "HND",
              "MEX",
              "NIC",
              "PAN",
              "PER",
              "PRY",
              "SLV",
              "URY",
              "VEN")


#GINI
for (i in 1:length(countries)) {
  name=paste0("gini_", countries[i])
  pais= paste0(countries[i])
  assign(name, wb(indicator = "SI.POV.GINI", country = pais, startdate = 2005, enddate = 2019))
}

gini= rbind(gini_ARG,
             gini_BOL,
             gini_BRA,
             gini_CHL,
             gini_COL,
             gini_CRI,
             gini_DOM,
             gini_ECU,
             gini_GTM,
             gini_HND,
             gini_MEX,
             gini_NIC,
             gini_PAN,
             gini_PER,
             gini_PRY,
             gini_SLV,
             gini_URY,
             gini_VEN)

gini$id= NA
gini$id= paste0(gini$iso3c,"_",gini$date)
gini= gini[,c(8,1,2,3,4,5,6,7)]
gini= gini[,c(1,4)]
names(gini)[2]= "gini"


#GDP per capita
for (i in 1:length(countries)) {
  name=paste0("gdp_", countries[i])
  pais= paste0(countries[i])
  assign(name, wb(indicator = "NY.GDP.PCAP.KD", country = pais, startdate = 2005, enddate = 2019))
}

gdp= rbind(gdp_ARG,
             gdp_BOL,
             gdp_BRA,
             gdp_CHL,
             gdp_COL,
             gdp_CRI,
             gdp_DOM,
             gdp_ECU,
             gdp_GTM,
             gdp_HND,
             gdp_MEX,
             gdp_NIC,
             gdp_PAN,
             gdp_PER,
             gdp_PRY,
             gdp_SLV,
             gdp_URY,
             gdp_VEN)


gdp$id= NA
gdp$id= paste0(gdp$iso3c,"_",gdp$date)
gdp= gdp[,c(8,1,2,3,4,5,6,7)]
gdp= gdp[,c(1,4)]
names(gdp)[2]= "gdp"

#Thousands of dollars
gdp$gdp = (gdp$gdp)/1000

```

+ Imputations

```{r Country level variables imputation, echo=TRUE, message=FALSE, warning=FALSE}
#Fusion of country variables databases
country_vars = full_join(gini, gdp) %>% 
  arrange(desc(id))

#Original databases
country_vars_original = country_vars
lapop_original = lapop

#Reallocation of data in missings
country_vars$gini[country_vars$id=="ARG_2018"] = country_vars$gini[country_vars$id=="ARG_2017"]
country_vars$gini[country_vars$id=="BOL_2010"] = country_vars$gini[country_vars$id=="BOL_2009"]
country_vars$gini[country_vars$id=="BOL_2018"] = country_vars$gini[country_vars$id=="BOL_2017"]
country_vars$gini[country_vars$id=="BRA_2010"] = country_vars$gini[country_vars$id=="BRA_2009"]
country_vars$gini[country_vars$id=="BRA_2018"] = country_vars$gini[country_vars$id=="BRA_2017"]
country_vars$gini[country_vars$id=="CHL_2008"] = country_vars$gini[country_vars$id=="CHL_2009"]
country_vars$gini[country_vars$id=="CHL_2010"] = country_vars$gini[country_vars$id=="CHL_2009"]
country_vars$gini[country_vars$id=="CHL_2012"] = country_vars$gini[country_vars$id=="CHL_2011"]
country_vars$gini[country_vars$id=="CHL_2014"] = country_vars$gini[country_vars$id=="CHL_2013"]
country_vars$gini[country_vars$id=="CHL_2016"] = country_vars$gini[country_vars$id=="CHL_2015"]
country_vars$gini[country_vars$id=="CHL_2018"] = country_vars$gini[country_vars$id=="CHL_2017"]
country_vars$gini[country_vars$id=="COL_2018"] = country_vars$gini[country_vars$id=="COL_2017"]
country_vars$gini[country_vars$id=="CRI_2018"] = country_vars$gini[country_vars$id=="CRI_2017"]
country_vars$gini[country_vars$id=="DOM_2018"] = country_vars$gini[country_vars$id=="DOM_2016"] #DOM_2018
country_vars$gini[country_vars$id=="ECU_2018"] = country_vars$gini[country_vars$id=="ECU_2017"]
country_vars$gini[country_vars$id=="HND_2018"] = country_vars$gini[country_vars$id=="HND_2017"]
country_vars$gini[country_vars$id=="MEX_2018"] = country_vars$gini[country_vars$id=="MEX_2016"] #MEX_2018
country_vars$gini[country_vars$id=="PAN_2018"] = country_vars$gini[country_vars$id=="PAN_2017"]
country_vars$gini[country_vars$id=="PER_2018"] = country_vars$gini[country_vars$id=="PER_2017"]
country_vars$gini[country_vars$id=="PRY_2018"] = country_vars$gini[country_vars$id=="PRY_2017"]
country_vars$gini[country_vars$id=="SLV_2018"] = country_vars$gini[country_vars$id=="SLV_2017"]
country_vars$gini[country_vars$id=="URY_2018"] = country_vars$gini[country_vars$id=="URY_2017"]

#ECLAC
country_vars$gini[country_vars$id=="GTM_2006"] = 55.8
country_vars$gini[country_vars$id=="GTM_2014"] = 53.5
country_vars$gini[country_vars$id=="NIC_2009"] = 46.3
country_vars$gini[country_vars$id=="NIC_2014"] = 49.5
country_vars$gini[country_vars$id=="VEN_2008"] = 37.9
country_vars$gini[country_vars$id=="VEN_2010"] = 36.4
country_vars$gini[country_vars$id=="VEN_2012"] = 38.4
country_vars$gini[country_vars$id=="VEN_2014"] = 37.8

#Linear interpolation
country_vars$country_wave1 = country_vars$id

country_vars1 = country_vars %>% 
  separate(country_wave1, c("country", "year"))

#GTM
country_vars_GTM = country_vars1 %>% 
  filter(country=="GTM" & year>=2006 & year<=2014) %>% 
  mutate(gini=na.approx(gini))

#NIC
country_vars_NIC = country_vars1 %>% 
  filter(country=="NIC" & year>=2009 & year<=2014) %>% 
  mutate(gini=na.approx(gini))

#VEN
country_vars_VEN = country_vars1 %>% 
  filter(country=="VEN" & year<=2014)

#Rbind
country_vars_else = country_vars1 %>% 
  filter(country!="GTM") %>% 
  filter(country!="NIC") %>% 
  filter(country!="VEN") 

country_vars = rbind(country_vars_else, country_vars_GTM, country_vars_NIC, country_vars_VEN) %>%
  rename(country_wave = id) %>% 
  mutate(year = as.numeric(year)) %>% 
  arrange(desc(country_wave))

rm(list=setdiff(ls(), c("lapop", "country_vars", "lapop_original", "country_vars_original", "gini", "gdp")))
```

```{r Country level variables, echo=TRUE, message=FALSE, warning=FALSE}
#Waves
country_vars_waves = filter(country_vars, year==2008 | year==2010 | year==2012 | year==2014 | year==2016 | year==2018)

#Means
country_vars_waves = country_vars_waves %>% 
  group_by(country) %>% 
  mutate(gini_mean = mean(gini, na.rm = T),
         gdp_mean = mean(gdp, na.rm = T)) %>% 
  ungroup()

#Differences
country_vars_waves = country_vars_waves %>% 
  mutate(gini_dif = gini - gini_mean,
         gdp_dif = gdp - gdp_mean)

#Welfare state
country_vars_waves = country_vars_waves %>% 
  mutate(welfare = case_when(
    (country == "BRA" | country == "CRI" | country == "PAN" | country == "MEX" | country == "URY") ~ "Proteccionist",
    (country == "ARG" | country == "CHL") ~ "Productivist",
    TRUE ~"informal-familialist"))

```

# Merge

```{r Merge individual and country variables, echo=TRUE, message=FALSE, warning=FALSE}
#Merge Level-1 & Lvele-2 variables
country_vars_waves = country_vars_waves %>% 
  select(country, year, country_wave, gini, gdp, 
         gini_mean, gdp_mean, gini_dif, gdp_dif, welfare)

lapop = lapop %>%
  left_join(country_vars_waves)

#Generate year_rec & country_wave_rec
lapop = lapop %>% 
  mutate(year_rec = case_when(year == 2008 ~ "08",
                              year == 2010 ~ "10",
                              year == 2012 ~ "12",
                              year == 2014 ~ "14",
                              year == 2016 ~ "16",
                              year == 2018 ~ "18"),
         country_wave_rec = paste0(country, sep ="_", year_rec))

#Country and year as factor
lapop = lapop %>% 
  mutate(country = as.factor(country),
         year = as.factor(year))

#Only cases with Level-2 variables
lapop_all = lapop
lapop = lapop %>% 
  filter_at(vars(gini:welfare), all_vars(!is.na(.)))

```

# Final database

```{r Save and Load Databases, echo=TRUE, message=FALSE, warning=FALSE}
#Save Databases paper
rm(list=setdiff(ls(), c("lapop", "lapop_all", "country_vars_waves")))

#Save final database
save.image(here("Input", "Proc", "LAPOP_Final.RData"))

#Detach package conflicted
detach("package:conflicted", unload = TRUE)

```
